---
title: 'R for Data Science: Explore'
author: "Frederick Solt"
date: "September 13, 2016"
output: 
    ioslides_presentation:
        smaller: true
---

<!-- ```{r setup, include=FALSE} -->
<!-- knitr::opts_chunk$set(echo = FALSE) -->
<!-- ``` -->

Explore: Visualize
===

## Agenda

- Grammer of Graphics

- Using ggplot2

This presentation is based on the [ggplot2 tutorial](https://github.com/jennybc/ggplot2-tutorial) written by Jennifer Bryan. 

## ggplot2

- `ggplot2` is a data visualization package

- created by Hadley Wickham, first released in 2005

- implements Leland Wilkinson's (1999) _Grammar of Graphics_ scheme

- a core part of the `tidyverse` collection of packages

## Grammar of Graphics

**Elements**

* `data`: The data that you want to visualise

* `aes`: Aesthetic mappings
describing how variables in the data are mapped to aesthetic attributes

    + horizontal position (`x`)
    + vertical position (`y`)
    + colour
    + size

## Grammar of Graphics

**Elements**

- `geoms`: Geometric objects that represent what you actually see on
the plot
    + points
    + lines
    + polygons
    + bars

## Grammar of Graphics

**Elements**

- `stats`: Statistics transformations
    + binning and counting observations to create a histogram,
    + summarising a 2d relationship with a linear model
    + `stats` are optional


## Grammar of Graphics

**Elements**

- `scales`: relate the data to the aesthetic

- `coord`: a coordinate system that describes how data coordinates are
mapped to the plane of the graphic.

- `facet`: a faceting specification describes how to break up the data into sets.

## Grammar of Graphics

**Layers**

A layer is composed of four parts:

* data and aesthetic mapping
* a statistical transformation (`stat`)
* a geometric object (`geom`)
* a position adjustment

A plot is constructed by adding layers to each other

## ggplot2

Load the library
```{r, echo=TRUE}
library(ggplot2)
```

Read data
```{r, echo=TRUE}
library(gapminder)
str(gapminder)
```

## Scatterplot
```{r}
ggplot(gapminder, aes(x = gdpPercap, y = lifeExp)) + 
    geom_point()
```

## Data transformation, quick and dirty
```{r}
ggplot(gapminder, aes(x = log10(gdpPercap), y = lifeExp)) +
    geom_point()
```

## Data transformation
```{r}
ggplot(gapminder, aes(x = gdpPercap, y = lifeExp)) + 
    geom_point() +
    scale_x_log10()
```

## Color by continent
```{r}
ggplot(gapminder, aes(x = gdpPercap, y = lifeExp, color = continent)) + 
    geom_point() +
    scale_x_log10()
```


## Transparency
```{r}
ggplot(gapminder, aes(x = gdpPercap, y = lifeExp)) + 
    geom_point(alpha = (1/3), size = 3) + 
    scale_x_log10() 

```


## Curves
```{r}
ggplot(gapminder, aes(x = gdpPercap, y = lifeExp)) + 
    geom_point(alpha = (1/3), size = 3) + 
    scale_x_log10() +
    geom_smooth()
```


## Curves

```{r}
ggplot(gapminder, aes(x = gdpPercap, y = lifeExp)) + 
    geom_point(alpha = (1/3), size = 3) + 
    scale_x_log10() +
    geom_smooth(lwd = 3, se = FALSE)
```


## Curves

```{r}
ggplot(gapminder, aes(x = gdpPercap, y = lifeExp)) + 
    geom_point(alpha = (1/3), size = 3) + 
    scale_x_log10() +
    geom_smooth(lwd = 3, se = FALSE, method = "lm")

```

## Curves

Return to continents:
```{r}
ggplot(gapminder, aes(x = gdpPercap, y = lifeExp, color = continent)) + 
    geom_point(alpha = (1/3), size = 3) + 
    scale_x_log10() +
    geom_smooth(lwd = 3, se = FALSE)
```


## Facets
```{r}
ggplot(gapminder, aes(x = gdpPercap, y = lifeExp, color = continent)) + 
    geom_point(alpha = (1/3), size = 3) + 
    scale_x_log10() +
    facet_wrap(~continent)

```

## Facets
```{r}
ggplot(gapminder, aes(x = gdpPercap, y = lifeExp, color = continent)) + 
    geom_point(alpha = (1/3), size = 3) + 
    scale_x_log10() +
    facet_wrap(~ continent) +
    geom_smooth(lwd = 2, se = FALSE)
```
