---
title: "R for Data Science: Exercises for Section Explore"
author: ["Garrett Grolemund", "Hadley Wickham"]
output: html_document
---

```{r, echo=FALSE}
library(ggplot2)
library(tibble)
```

## Visualize

### 3.2.1 Exercises

1.  Run `ggplot(data = mpg)` what do you see?

1.  What does the `drv` variable describe?  Read the help for `?mpg` to find
    out.
     
1.  Make a scatterplot of `hwy` vs `cyl`.

1.  What happens if you make a scatterplot of `class` vs `drv`. Why is
    the plot not useful?

### 3.3.1 Exercises

1.  What's gone wrong with this code? Why are the points not blue?

    ```{r}
    ggplot(data = mpg) + 
      geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
    ```
    
1.  Which variables in `mpg` are categorical? Which variables are continuous? 
    (Hint: type `?mpg` to read the documentation for the dataset). How
    can you see this information when you run `mpg`?

1.  Map a continuous variable to `color`, `size`, and `shape`. How do
    these aesthetics behave differently for categorical vs. continuous
    variables? 
    
1.  What happens if you map the same variable across multiple aesthetics? 
    What happens if you map different variables across multiple aesthetics?
  
1.  What does the `stroke` aesthetic do? What shapes does it work with?
    (Hint: use `?geom_point`)
    
1.  What happens if you set an aesthetic to something other than a variable 
    name, like `displ < 5`?

### 3.5.1 Exercises

1.  What happens if you facet on a continuous variable?

1.  What do the empty cells in plot with `facet_grid(drv ~ cyl)` mean?
    How do they relate to this plot?
    
    ```{r, eval = FALSE}
    ggplot(data = mpg) + 
      geom_point(mapping = aes(x = drv, y = cyl))
    ```

1.  What plots does the following code make? What does `.` do?

    ```{r eval = FALSE}
    ggplot(data = mpg) + 
      geom_point(mapping = aes(x = displ, y = hwy)) +
      facet_grid(drv ~ .)
    
    ggplot(data = mpg) + 
      geom_point(mapping = aes(x = displ, y = hwy)) +
      facet_grid(. ~ cyl)
    ```

1.  Take the first faceted plot in this section:

    ```{r, eval = FALSE}
    ggplot(data = mpg) + 
      geom_point(mapping = aes(x = displ, y = hwy)) + 
      facet_wrap(~ class, nrow = 2)
    ```
    
    What are the advantages to using faceting instead of the colour aesthetic?
    What are the disadvantages? How might the balance change if you had a 
    larger dataset?
    
1.  Read `?facet_wrap`. What does `nrow` do? What does `ncol` do? What other
    options control the layout of the individual panels? Why doesn't
    `facet_grid()` have `nrow` and `ncol` variables?

1.  When using `facet_grid()` you should usually put the variable with more
    unique levels in the columns. Why?

### 3.6.1 Exercises

1.  What geom would you use to draw a line chart? A boxplot? 
    A histogram? An area chart?

1.  Run this code in your head and predict what the output will look like.
    Then, run the code in R and check your predictions.
    
    ```{r, eval = FALSE}
    ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
      geom_point() + 
      geom_smooth(se = FALSE)
    ```
1.  What does `show.legend = FALSE` do?  What happens if you remove it?  
    Why do you think I used it in the example above.

1.  What does the `se` argument to `geom_smooth()` do?


1.  Will these two graphs look different? Why/why not?

    ```{r, eval = FALSE}
    ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
      geom_point() + 
      geom_smooth()
    
    ggplot() + 
      geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
      geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
    ```

1.  Recreate the R code necessary to generate the following graphs.
    
    ```{r echo = FALSE, fig.width = 3, out.width = "50%", fig.align = "default", message = FALSE}
    ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
      geom_point() + 
      geom_smooth(aes(group = drv), se = FALSE)
    ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
      geom_smooth(aes(group = drv), se = FALSE) +
      geom_point()
    ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
      geom_point() + 
      geom_smooth(se = FALSE)
    ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
      geom_point(aes(color = drv)) + 
      geom_smooth(se = FALSE)
    ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
      geom_point(aes(color = drv)) +
      geom_smooth(aes(linetype = drv), se = FALSE)
    ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
      geom_point(size = 4, colour = "white") + 
      geom_point(aes(colour = drv))
    ```

### 3.7.1 Exercises
    
1.  In our proportion barchart, we need to set `group = 1`. Why?  In other
    words, why is this graph not useful?
    
    ```{r, eval = FALSE}
    ggplot(data = diamonds) + 
      geom_bar(mapping = aes(x = cut, y = ..prop..))
    ```
    
1.  How do you find out the default stat associated with a geom?

### 3.8.1 Exercises

1.  What is the problem with this plot? How could you improve it?

    ```{r}
    ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
      geom_point()
    ```

1.  Compare and contrast `geom_jitter()` with `geom_count()`.

1.  What's the default position adjustment for `geom_boxplot()`? Create
    a visualisation of the `mpg` dataset that demonstrates it.

### 3.8.1 Exercises
    
1.  Turn a stacked bar chart into a pie chart using `coord_polar()`.

1.  What does `labs()` do? Read the documentation.

1.  What's the difference between `coord_quickmap()` and `coord_map()`?

1.  What does the plot below tell you about the relationship between city
    and highway mpg? Why is `coord_fixed()` important? What does 
    `geom_abline()` do?
    
    ```{r, fig.asp = 1, out.width = "50%"}
    ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
      geom_point() + 
      geom_abline() +
      coord_fixed()
    ```


## Workflow: basics

### 4.4 Practice

1.  Why does this code not work?

    ```{r, error = TRUE}
    my_variable <- 10
    my_varıable
    ```
    
    Look carefully! (This may seem like an exercise in pointlessness, but
    training your brain to notice even the tiniest difference will pay off
    when programming.)
    
1.  Tweak the each of the following R commands so that they run correctly:

    ```{r, eval = FALSE}
    library(ggplot2)
    library(dplyr)
    
    ggplot(dota = mpg) + 
      geom_point(mapping = aes(x = displ, y = hwy))
    
    mpg %>% 
      fliter(cyl = 8)
  
    diamond %>% 
      filter(carat > 3)
    ```
    
1.  Press Alt + Shift + K. What happens? How can you get to the same place
    using the menus?


## Data transformation

### 5.2.4 Exercises

1.  Find all flights that

    1. Had an arrival delay of two or more hours.
    1. Flew to Houston (`IAH` or `HOU`)
    1. Were operated by United, American, or Delta
    1. Departed in summer (July, August, and September)
    1. Arrived more than two hours late, but didn't leave late
    1. Were delayed by at least an hour, but made up over 30 minutes in flight
    1. Departed between midnight and 6am (inclusive)

1.  Another useful dplyr filtering helper is `between()`. What does it do?
    Can you use it to simplify the code needed to answer the previous 
    challenges?

1.  How many flights have a missing `dep_time`? What other variables are 
    missing? What might these rows represent?

1.  Why is `NA ^ 0` not missing? Why is `NA | TRUE` not missing?
    Why is `FALSE & NA` not missing? Can you figure out the general
    rule?  (`NA * 0` is a tricky counterexample!)

### 5.3.1 Exercises

1.  How could you use `arrange()` to sort all missing values to the start?
    (Hint: use `is.na()`).
    
1.  Sort `flights` to find the most delayed flights. Find the flights that
    left earliest.

1.  Sort `flights` to find the fastest flights.

1.  Which flights travelled the longest? Which travelled the shortest?

### 5.4.1 Exercises

1.  Brainstorm as many ways as possible to select `dep_time`, `dep_delay`,
    `arr_time`, and `arr_delay` from `flights`.
    
1.  What happens if you include the name of a variable multiple times in
    a `select()` call?
  
1.  What does the `one_of()` function do? Why might it be helpful in conjunction
    with this vector?
    
### 5.5.2 Exercises

1.  Currently `dep_time` and `sched_dep_time` are convenient to look at, but
    hard to compute with because they're not really continuous numbers. 
    Convert them to a more convenient representation of number of minutes
    since midnight.
    
1.  Compare `airtime` with `arr_time - dep_time`. What do you expect to see?
    What do you see? What do you need to do to fix it?
    
1.  Compare `dep_time`, `sched_dep_time`, and `dep_delay`. How would you
    expect those three numbers to be related?

1.  Find the 10 most delayed flights using a ranking function. How do you want 
    to handle ties? Carefully read the documentation for `min_rank()`.

1.  What does `1:3 + 1:10` return? Why?

1.  What trigonometric functions does R provide?

### 5.6.7 Exercises

1.  Brainstorm at least 5 different ways to assess the typical delay 
    characteristics of a group of flights. Consider the following scenarios:
    
    * A flight is 15 minutes early 50% of the time, and 15 minutes late 50% of 
      the time.
      
    * A flight is always 10 minutes late.

    * A flight is 30 minutes early 50% of the time, and 30 minutes late 50% of 
      the time.
      
    * 99% of the time a flight is on time. 1% of the time it's 2 hours late.
    
    Which is more important: arrival delay or departure delay?

1.  Come up with another approach that will give you the same output as 
    `not_cancelled %>% count(dest)` and 
    `not_cancelled %>% count(tailnum, wt = distance)` (without using 
    `count()`).

1.  Our definition of cancelled flights (`!is.na(dep_delay) & !is.na(arr_delay)`
    ) is slightly suboptimal. Why? Which is the most important column?

1.  Look at the number of cancelled flights per day. Is there a pattern?
    Is the proportion of cancelled flights related to the average delay?

1.  Which carrier has the worst delays? Challenge: can you disentangle the
    effects of bad airports vs. bad carriers? Why/why not? (Hint: think about
    `flights %>% group_by(carrier, dest) %>% summarise(n())`)

1.  For each plane, count the number of flights before the first delay 
    of greater than 1 hour.

1.  What does the `sort` argument to `count()` do. When might you use it?

### 5.7.1 Exercises

1.  Refer back to the table of useful mutate and filtering functions. 
    Describe how each operation changes when you combine it with grouping.

1.  Which plane (`tailnum`) has the worst on-time record?

1.  What time of day should you fly if you want to avoid delays as much
    as possible?

  
## Exploratory Data Analysis

### 7.3.4 Exercises

1.  Explore the distribution of each of the `x`, `y`, and `z` variables 
    in `diamonds`. What do you learn? Think about a diamond and how you
    might decide which dimension is the length, width, and depth.

1.  Explore the distribution of `price`. Do you discover anything unusual
    or surprising? (Hint: Carefully think about the `binwidth` and make sure
    you try a wide range of values.)

1.  How many diamonds are 0.99 carat? How many are 1 carat? What
    do you think is the cause of the difference?
    
1.  Compare and contrast `coord_cartesian()` vs `xlim()` or `ylim()` when
    zooming in on a histogram. What happens if you leave `binwidth` unset?
    What happens if you try and zoom so only half a bar shows?
    
1.  Delays are typically temporally correlated: even once the problem that
    caused the initial delay has been resolved, later flights are delayed 
    to allow earlier flights to leave. Using `lag()` explore how the delay
    of a flight is related to the delay of the immediately preceding flight.
    
1.  Look at each destination. Can you find flights that are suspiciously
    fast? (i.e. flights that represent a potential data entry error). Compute
    the air time a flight relative to the shortest flight to that destination.
    Which flights were most delayed in the air?
    
1.  Find all destinations that are flown by at least two carriers. Use that
    information to rank the carriers.

### 7.4.1 Exercises

1.  What happens to missing values in a histogram?  What happens to missing
    values in bar chart? Why is there a difference?

1.  What does `na.rm = TRUE` do in `mean()` and `sum()`?

#### 7.5.1.1 Exercises

1.  Use what you've learned to improve the visualisation of the departure times
    of cancelled vs. non-cancelled flights.

1.  What variable in the diamonds dataset is most important for predicting
    the price of a diamond? How is that variable correlated with cut?
    Why does the combination of those two relationships lead to lower quality
    diamonds being more expensive?

1.  Install the ggstance package, and create a horizontal boxplot.
    How does this compare to using `coord_flip()`?

1.  One problem with boxplots is that they were developed in an era of 
    much smaller datasets and tend to display a prohibitively large
    number of "outlying values". One approach to remedy this problem is
    the letter value plot. Install the lvplot package, and try using
    `geom_lv()` to display the distribution of price vs cut. What
    do you learn? How do you interpret the plots?

1.  Compare and contrast `geom_violin()` with a facetted `geom_histogram()`,
    or a coloured `geom_freqpoly()`. What are the pros and cons of each 
    method?

1.  If you have a small dataset, it's sometimes useful to use `geom_jitter()`
    to see the relationship between a continuous and categorical variable.
    The ggbeeswarm package provides a number of methods similar to 
    `geom_jitter()`. List them and briefly describe what each one does.
    
#### 7.5.2.1 Exercises

1.  How could you rescale the count dataset above to more clearly show
    the distribution of cut within colour, or colour within cut?

1.  Use `geom_tile()` together with dplyr to explore how average flight
    delays vary by destination and month of year.  What makes the 
    plot difficult to read? How could you improve it?

1.  Why is it slightly better to use `aes(x = color, y = cut)` rather
    than `aes(x = cut, y = color)` in the example above?

#### 7.5.3.1 Exercises

1.  Instead of summarising the conditional distribution with a boxplot, you
    could use a frequency polygon. What do you need to consider when using
    `cut_width()` vs `cut_number()`? How does that impact a visualisation of
    the 2d distribution of `carat` and `price`?

1.  Visualise the distribution of carat, partitioned by price.

1.  How does the price distribution of very large diamonds compare to small 
    diamonds. Is it as you expect, or does it surprise you?
    
1.  Combine two of the techniques you've learned to visualise the 
    combined distribution of cut, carat, and price.

1. Two dimensional plots reveal outliers that are not visible in one 
   dimensional plots. For example, some points in the plot below have an 
   unusual combination of `x` and `y` values, which makes the points outliers 
   even though their `x` and `y` values appear normal when examined separately.
  
    ```{r, dev = "png"}
    ggplot(data = diamonds) +
      geom_point(mapping = aes(x = x, y = y)) +
      coord_cartesian(xlim = c(4, 11), ylim = c(4, 11))
    ```
    
    Why is a scatterplot a better display than a binned plot for this case?
